---
title: 'Main Takeaways'
output:
  distill::distill_article:
    toc:false:
      theme: theme.css
favicon: "images/favicon.ico"
github-repo: childpovertyactionlab/dpd-incident-report
---

```{r Set-Up Block, include = FALSE}
#### Libraries to load #####
library(tidyverse)
library(sf)
library(leaflet)
#library(mapboxapi)
library(lubridate)
library(rio)
#library(tidycensus)
library(htmltools)
#library(arcgisbinding)
library(reactable)
library(CPALtools)

#datalib <- "C:/Users/micha/CPAL Dropbox/" # Michael Laptop
datalib <- "E:/CPAL Dropbox/" # Michael Desktop

#### Import Arrest Data 
arrests <- import(paste0(datalib, "Data Library/City of Dallas/04_Public Safety/Dallas Police/Data/Arrests/Processed Data/Dallas Police Arrests - All Arrests.csv")) %>%
  select(incidentnum, arrestnumber, ararrestdate, ararresttime, arpremises)

#### Defining vgc and firearms #####
firearm <- c("Handgun", "Rifle","Firearm (Type Not Stated)", "Assault Weapon", "Unknown Type Gun", "Other/Unknown Gun", "Shotgun", "Other Gun", "Other Firearm")
vgc_type <- c("AGG ASSAULT - NFV", "MURDER & NONNEGLIGENT MANSLAUGHTER", "NEGLIGENT MANSLAUGHTER")

#### Importing Dallas Open Data portals and filtering into data frames #####
dpdIncidents <- st_read(paste0(datalib, "Data Library/City of Dallas/04_Public Safety/Dallas Police/Data/Incidents/Processed Data/ODP - Dallas Police Incidents.gpkg"), layer = "Group A") %>%
  left_join(., arrests) %>%
  mutate(Date = as.Date(Date),
         Time = lubridate::hm(Time),
         Year = year(Date),
         Month = month(Date),
         Day = day(Date),
         division = str_to_title(division),
         crime_category = as.factor(nibrs_crime_category),
         crime = as.factor(nibrs_crime),
         Weekday = wday(Date),
         DateTime = format(paste(Date, Time), format="%Y-%m-%d %H:%M:%S"),
         hour = hour(Time),
         vgc_flag = ifelse(nibrs_crime %in% vgc_type & Firearm == 1, 1, 0),
         ToD = ifelse(hour >= 3 & hour < 7, "Early Morning",
                      ifelse(hour >= 7 & hour < 13, "Morning",
                             ifelse(hour >= 13 & hour < 18, "Afternoon",
                                    ifelse(hour >= 18 & hour < 23, "Evening",
                                           ifelse(hour >= 23 | hour < 3, "Night", "NA")
                                           )
                                    )
                             )
                      )
         ) %>%
  mutate(crime_title = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", "Murder & Nonnegligent Manslaughter",
                              ifelse(crime == "VIOLENT GUN CRIME", "Violent Gun Crime",
                                     ifelse(crime == "AGG ASSAULT - NFV", "Aggravated Assault (NFV)",
                                            ifelse(crime == "ROBBERY-BUSINESS", "Robbery (Business)",
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", "Robbery (Individual)",
                                                          ifelse(crime == "BURGLARY-RESIDENCE", "Burglary (Residence)",
                                                                 ifelse(crime == "BURGLARY-BUSINESS", "Burglary (Business)",
                                                                        "ERROR"
                              )))))))) %>%
  st_transform(crs = 4269) %>%
  select(incidentnum:premise, incident_address:time1, reporteddate, involvement:ro1name, status, mo, offensecode:nibrs_type, zip_code:geocoded_column.longitude, taag:compsex, weaponused, tot_victims:crime_title) %>%
  filter(crime %in% c("MURDER & NONNEGLIGENT MANSLAUGHTER", "AGG ASSAULT - NFV", "ROBBERY-BUSINESS", "ROBBERY-INDIVIDUAL", "BURGLARY-RESIDENCE", "BURGLARY-BUSINESS")) %>%
    mutate(ArDate = as.Date(ararrestdate),
         ArLength = time_length(ArDate-Date, unit = "days"),
         ArOccur = ifelse(is.na(arrestnumber), "NO", "YES"))


#unique(dpdIncidents$crime)


# generate dataframes based on various time cuts needed for visuals
currentmonth <- floor_date(today(), unit = "month")
last30 <- currentmonth-months(1)
lastweek <- as.Date(max(dpdIncidents$date1))-days(7)
twoweeks <- as.Date(max(dpdIncidents$date1))-days(14)

# generate vgc data frames
# what need is there for these?
vgc5year <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= currentmonth - years(5))

vgc12months <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= currentmonth - years(1))

vgc30days <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= last30 & Date < currentmonth)

vgc7days <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= lastweek)

dpdMonth <- dpdIncidents %>%
  filter(Date >= last30 & Date < currentmonth)

dpdWeek <- dpdIncidents %>%
  filter(Date >= lastweek)

arMonth <- dpdIncidents %>%
  filter(ArDate >= last30 & ArDate < currentmonth)

#### Import tenth mile grid of the City of Dallas #####
grid_incidents <- st_read("data/dallas_tenthmilegrid.geojson") %>%
  st_transform(crs = 4269) %>%
  mutate(vgc_5year = lengths(st_intersects(., vgc5year)),
         vgc_12months = lengths(st_intersects(., vgc12months)),
         vgc_5yearAvg = vgc_5year/5,
         vgc_perch = (vgc_12months-vgc_5yearAvg)/vgc_5yearAvg,
         vgc_perch = ifelse(vgc_perch == "NaN", NA, vgc_perch))

#### Import  Dallas boundaries #####
DallasBoundary <- st_read("data/dallas_simpleboundary.geojson") %>%
  st_transform(crs = 4269)

dpd_beats <- st_read("data/dpd_beats.geojson") %>%
  st_transform(crs = 4269)

dpd_divisions <- st_read("data/dpd_divisions.geojson") %>%
  st_transform(crs = 4269)

thisyear <- year(last30)
lastyear <- year(last30-years(1))
thismonth <- month(last30)
lastmonth <- month(last30 - month(1))

month_comp <- dpdIncidents %>%
  filter(crime %in% vgc_type & Firearm == 1) %>%
  filter((Year == thisyear & Month == thismonth) | #most recent complete month
           (Year == lastyear & Month == thismonth) | #year prior but same month
           (Year == thisyear & Month == lastmonth)) %>% #last month
  st_drop_geometry(.) %>%
  group_by(Year, Month) %>%
  summarise(count = n())

month_01 <- month_comp %>%
  filter(Year == thisyear & Month == month(last30))

month_02 <- month_comp %>%
  filter(Year == thisyear & Month == lastmonth)

month_03 <- month_comp %>%
  filter(Year == lastyear & Month == thismonth)

month_04 <- dpdIncidents %>%
  filter((Year == thisyear & Month == thismonth)) %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month) %>%
  summarise(count = n())

week_comp <- dpdIncidents %>%
  filter(crime %in% vgc_type & Firearm == 1) %>%
  mutate(weekFlag = ifelse(Date >= lastweek, "This Week",
                           ifelse(Date >= twoweeks & Date < lastweek, "Last Week",
                                  "IGNORE"))) %>%
  filter(weekFlag != "IGNORE") %>%
  st_drop_geometry(.) %>%
  group_by(weekFlag) %>%
  summarise(count = n())

week_01 <- week_comp %>%
  filter(weekFlag == "This Week")

week_02 <- week_comp %>%
  filter(weekFlag == "Last Week")

week_03 <- dpdIncidents %>%
  mutate(weekFlag = ifelse(Date >= lastweek, "This Week",
                           ifelse(Date >= twoweeks & Date < lastweek, "Last Week",
                                  "IGNORE"))) %>%
  filter(weekFlag == "This Week") %>%
  st_drop_geometry(.) %>%
  group_by(weekFlag) %>%
  summarise(count = n())
```

```{r, High Risk Cells, include = FALSE}
#### Import High Risk RTM Cells from most recent models
#st_layers(paste0(datalib, "Safe Surroundings/04_Projects/RTM 2022/2022_Q1_RTM/Data/2022_Q1_RTM_Outputs.gpkg"))
hr_directory <- paste0(datalib, "Safe Surroundings/04_Projects/RTM 2022/2022_Q1_RTM/Data/2022_Q1_RTM_Outputs.gpkg")

SE_HR <- st_read(hr_directory, layer = "SE RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)
SC_HR <- st_read(hr_directory, layer = "SC RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)
C_HR <- st_read(hr_directory, layer = "C RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)
NC_HR <- st_read(hr_directory, layer = "NC RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)
NE_HR <- st_read(hr_directory, layer = "NE RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)
NW_HR <- st_read(hr_directory, layer = "NW RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)
SW_HR <- st_read(hr_directory, layer = "SW RTM RRSz") %>%
  st_transform(crs = 4269) %>%
  filter(highest == 1)

SC_Mark <- vgc30days %>%
  filter(division == "South Central") %>%
  .[SC_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

SE_Mark <- vgc30days %>%
  filter(division == "Southeast") %>%
  .[SE_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

NE_Mark <- vgc30days %>%
  filter(division == "Northeast") %>%
  .[NE_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

NW_Mark <- vgc30days %>%
  filter(division == "Northwest") %>%
  .[NW_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

C_Mark <- vgc30days %>%
  filter(division == "Central") %>%
  .[C_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

SW_Mark <- vgc30days %>%
  filter(division == "Southwest") %>%
  .[SW_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

NC_Mark <- vgc30days %>%
  filter(division == "North Central") %>%
  .[NC_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

vgc30days <- bind_rows(SC_Mark, SE_Mark, NE_Mark, NW_Mark, C_Mark, SW_Mark, NC_Mark) %>%
  left_join(vgc30days, .)

Division_VGC <- vgc30days %>%
  st_drop_geometry(.) %>%
  filter(vgc_flag == 1) %>%
  group_by(division) %>%
  summarise(count = n()) %>%
  mutate(perc = count/sum(count))

maxDivision <- filter(Division_VGC, perc == max(perc))

vgc_hr <- vgc30days %>%
  st_drop_geometry(.) %>%
  filter(vgc_flag == 1) %>%
  group_by(HighRisk) %>%
  summarise(count = n()) %>%
  filter(HighRisk == TRUE)
```

# `r paste(month(thismonth, label = TRUE, abbr = FALSE), thisyear)`

## Last Month Summary

The violent crime rate for **`r paste(month(thismonth, label = TRUE, abbr = FALSE), thisyear)`** was **`r round(month_01$count/(1304379/10000), digits = 3)`** per 10,000 residents.

In the month of **`r paste(month(thismonth, label = TRUE, abbr = FALSE), thisyear)`** there were a total of **`r month_01$count`** violent gun crimes across the city.

```{r, Monthly VGC Column Chart, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
month_tbl <- dpdIncidents %>%
  st_drop_geometry(.) %>%
  filter(crime %in% vgc_type & Firearm == 1) %>%
  filter(Date >= as.Date("2018-01-01")) %>%
  group_by(Year, Month) %>%
  summarize(vgcCount = n()) %>%
  ungroup(.) %>%
  mutate(Date = as.Date(paste0(Year, "-", Month, "-01")),
         monthAvg = zoo::rollmean(vgcCount, k = 3, fill = NA))

month_tbl %>%
  ggplot(aes(x = Date, y = vgcCount)) +
  geom_col(fill = "#595959",
           alpha = 0.5) +
  geom_line(aes(y = monthAvg),
            color = "#ffffff",
            size = 2) +
  geom_line(aes(y = monthAvg),
            color = "#e98816",
            size = 1) +
  labs(
    title = "Violent Gun Crime by Month",
    subtitle = "Relative to 3 Month Average",
    x = "",
    y = "",
    color = 'VARIABLE'
  ) +

  theme_cpal()

```

This is **`r ifelse(month_01$count > month_02$count, "an increase", "a decrease")` of `r paste0(round(((month_01$count-month_02$count)/month_02$count)*100, digits = 1), "%")`** from **`r paste(month(lastmonth, label = TRUE, abbr = FALSE), thisyear)`**, and **`r ifelse(month_01$count > month_03$count, "an increase", "a decrease")`** of **`r paste0(round(((month_01$count-month_03$count)/month_03$count)*100, digits = 1), "%")`** from **`r paste(month(thismonth, label = TRUE, abbr = FALSE), lastyear)`**.

There were **`r nrow(filter(vgc30days, tot_victims >= 4))`** incidents with 4 or more victims across the city last month.

**`r paste0(round(max(Division_VGC$perc)*100, digit = 1), "%")`** of all violent gun crimes in **`r paste(month(thismonth, label = TRUE, abbr = FALSE), thisyear)`** occurred in **`r maxDivision$division`** Division

```{r, Pie Chart Division, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
pie_tbl <-vgc30days %>%
  st_drop_geometry(.) %>%
  group_by(division) %>%
  summarize(vgcCount = n()) %>%
  mutate(vgcPercent = round(vgcCount/sum(vgcCount), digits = 3)*100) %>%
  arrange(desc(vgcPercent)) %>%
  mutate(ypos = cumsum(vgcPercent)- 0.5*vgcPercent,
         division = ifelse(division == "North Central", "NC",
                           ifelse(division == "Northwest", "NW",
                                  ifelse(division == "Northeast", "NE",
                                         ifelse(division == "South Central", "SC",
                                                ifelse(division == "Southwest", "SW",
                                                       ifelse(division == "Southeast", "SE",
                                                              ifelse(division == "Central", "C", "ERROR"
                                  ))))))))

pie_tbl %>%
  ggplot(aes(x = "", y = vgcPercent, fill = reorder(division, vgcPercent))) +
  geom_bar(stat = "identity", width = 1, color = "#ffffff", size = 1) +
  scale_fill_manual(values = pal_cpalfac) +
  coord_polar("y", start = 0) +
  theme_cpal() +
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        ) +
  geom_text(aes(x = 1.2, 
                y = ypos, 
                label = paste0(division, "\n",
                               vgcPercent, "%")), 
            color = "white", 
            fontface = "bold",
            size=12) +
  labs(
    title = "Percent of Violent Gun Crimes by Division",
    subtitle = "",
    x = "",
    y = "",
    color = 'VARIABLE'
  )
```

**`r paste0(round(vgc_hr$count/month_01$count*100, digit = 1), "%", " (", vgc_hr$count, ")")`** of Violent Gun Crimes occurred within a cell identified as "High Risk" by our most recent Risk Terrain Modeling Analysis.

### Arrests
```{r, Arrests, include = FALSE}
ArVGC <- dpdMonth %>%
  st_drop_geometry(.) %>%
  filter(vgc_flag == 1) %>%
  group_by(ArOccur) %>%
  summarize(count = n(),
            ArMedian = median(ArLength))

ArSameDay <- dpdMonth %>%
  st_drop_geometry(.) %>%
  filter(vgc_flag == 1) %>%
  filter(ArLength == 0) %>%
  group_by(ArOccur) %>%
  summarize(count = n())

ArYes <- ArVGC %>%
  filter(ArOccur == "YES")
```

In the last month there were a total of **`r paste0(ArYes$count)`** arrests for Violent Gun Crimes across the City of Dallas, which accounted for **`r paste0(round(ArYes$count/sum(ArVGC$count), digits = 3)*100, "%")`** of all Violent Gun Crime incidents.

`r paste0(ArSameDay$count)` arrests occurred on the same day of the incident.

## Last Week Summary

For the week of **`r paste(format(lastweek, format = "%B %d, %Y"))`** there was a total of **`r week_01$count`** violent gun crimes across the city. This is a difference of **`r week_01$count - week_02$count`** from the prior week.

```{r, Weekly VGC Column Chart, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
week_tbl <- dpdIncidents %>%
  st_drop_geometry(.) %>%
  filter(crime %in% vgc_type & Firearm == 1) %>%
  filter(Date >= as.Date("2018-01-01")) %>%
  group_by(Date) %>%
  summarize(vgcCount = n()) %>%
  mutate(sevenAvg = zoo::rollmean(vgcCount, k = 28, fill = NA))

week_tbl %>%
  ggplot(aes(x = Date, y = vgcCount)) +
  geom_col(fill = "#595959",
           alpha = 0.5) +
  geom_line(aes(y = sevenAvg),
            color = "#ffffff",
            size = 1.5) +
  geom_line(aes(y = sevenAvg),
            color = "#e98816",
            size = 1) +
  scale_y_continuous(limits = c(0, 15)) +
  labs(
    title = "Violent Gun Crime by Week",
    subtitle = "Relative to 28 Day Average",
    x = "",
    y = "",
    color = 'VARIABLE'
  ) +

  theme_cpal()
```

There were **`r nrow(filter(vgc7days, tot_victims >= 4))`** incidents with 4 or more victims across the city last week.

The reported U.S. violent crime rate includes murder, rape and sexual assault, robbery, and assault
