---
title: 'Insights This Month'
output:
  distill::distill_article:
    toc:false:
      theme: theme.css
favicon: "images/favicon.ico"
github-repo: childpovertyactionlab/dpd-incident-report
---

## Select Group A Incidents Overview {-}

The table below contains an overview of selected group A incidents for the most recent complete month (`r paste(month(thismonth, label = TRUE, abbr = FALSE), thisyear)`). Data focuses on the count of incidents within the full month compared to the same month for the previous year and a three year average for the specified month. 

* Note if Poisson E value is +/-3 then there is a significant difference between incident counts across time periods.

```{r, Incidents Table, echo = FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=10, layout = "l-body-outset"}
incident_nibrs <- dpdIncidents %>%
  filter(crime %in% c("AGG ASSAULT - NFV", "ROBBERY-BUSINESS", "ROBBERY-INDIVIDUAL", "BURGLARY-RESIDENCE", "BURGLARY-BUSINESS", "MURDER & NONNEGLIGENT MANSLAUGHTER")) %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month, crime) %>%
  summarize(count = n())

incident_vgc <- dpdIncidents %>%
  filter(vgc_flag == 1 & crime %in% vgc_type) %>%
  mutate(crime = "VIOLENT GUN CRIME") %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month, crime) %>%
  summarize(count = n())

incident_table <- full_join(incident_nibrs, incident_vgc) %>%
  mutate(Date = as.Date(paste0(Year, "-", Month, "-01"))) %>%
  filter(Year > 2016,
         Date <= currentmonth) %>%
  mutate(crime_title = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", "Murder & Nonnegligent Manslaughter",
                              ifelse(crime == "VIOLENT GUN CRIME", "Violent Gun Crime",
                                     ifelse(crime == "AGG ASSAULT - NFV", "Aggravated Assault (NFV)",
                                            ifelse(crime == "ROBBERY-BUSINESS", "Robbery (Business)",
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", "Robbery (Individual)",
                                                          ifelse(crime == "BURGLARY-RESIDENCE", "Burglary (Residence)",
                                                                 ifelse(crime == "BURGLARY-BUSINESS", "Burglary (Business)",
                                                                        "ERROR"
                              ))))))))

recent_month <- month(last30)

report_table <- incident_table %>%
  filter(Month == recent_month) %>%
  ungroup(.) %>%
  select(-Month, -Date) %>%
  mutate(
         crime = factor(crime, levels = c("MURDER & NONNEGLIGENT MANSLAUGHTER", "VIOLENT GUN CRIME", "AGG ASSAULT - NFV", "ROBBERY-BUSINESS", "ROBBERY-INDIVIDUAL", "BURGLARY-RESIDENCE", "BURGLARY-BUSINESS")),
         crime_title = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", "Murder & Nonnegligent Manslaughter",
                              ifelse(crime == "VIOLENT GUN CRIME", "Violent Gun Crime",
                                     ifelse(crime == "AGG ASSAULT - NFV", "Aggravated Assault (NFV)",
                                            ifelse(crime == "ROBBERY-BUSINESS", "Robbery (Business)",
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", "Robbery (Individual)",
                                                          ifelse(crime == "BURGLARY-RESIDENCE", "Burglary (Residence)",
                                                                 ifelse(crime == "BURGLARY-BUSINESS", "Burglary (Business)",
                                                                        "ERROR"
                              ))))))),
         crime_order = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", 1,
                              ifelse(crime == "VIOLENT GUN CRIME", 2,
                                     ifelse(crime == "AGG ASSAULT - NFV", 3,
                                            ifelse(crime == "ROBBERY-BUSINESS", 4,
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", 5,
                                                          ifelse(crime == "BURGLARY-RESIDENCE", 6,
                                                                 ifelse(crime == "BURGLARY-BUSINESS", 7,
                                                                        "ERROR"
                              )))))))) %>%
  select(-crime) %>%
  relocate(crime_order, crime_title)

##### Table
reactable(report_table,
          compact = TRUE,
          class = "cpal-tbl",
          defaultSorted = c("crime_order"),
          columns = list(
            crime_order = colDef(
              name = "Order",
              show = FALSE),
            crime_title = colDef(
              name = "NIBRS"),
            dpd_2021 = colDef(
              name = "2021"),
            dpd_2022 = colDef(
              name = "2022"),
            poissonYtY = colDef(
              name = "Poisson E",
              defaultSortOrder = "desc",
              align = "center"
              ),
            Year2Year = colDef(
              name = "1-Year",
              defaultSortOrder = "desc",
              align = "center"
              ),
            Avg3Year = colDef(
              name = "3-Year",
              defaultSortOrder = "desc")
            ),
          theme = reactableTheme(
      headerStyle = list(
        "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
        "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
        borderColor = "#555"
        )
      ),
      columnGroups = list(
            colGroup(name = "Percent Change", columns = c("Year2Year")),
            colGroup(name = "Average", columns = c("Avg3Year"))
          )
            )
```

```{r, Incident Line Graph, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
#### Plot Time Series by Month #####
incident_table %>%
  filter(Date != currentmonth) %>%
  filter(crime_title != "Murder & Nonnegligent Manslaughter") %>%
  filter(Date >= as.Date("2018-01-01")) %>%
  mutate(YearMonth = as.Date(paste0(Year, "-", Month, "-01"))) %>%
  group_by(crime_title) %>%
  arrange(YearMonth) %>%
  mutate(Avg3Month = zoo::rollmean(count, k = 3, fill = NA)) %>%
  ungroup(.) %>%
  ggplot(
    aes(
      x = YearMonth,
      y = count,
      group = crime_title,
      color = crime_title
    )
  ) +
  geom_col(fill = "#595959",
           color = "#595959",
           alpha = 0.5) +
  geom_line(size=2, alpha = 0.9, color = "#ffffff", stat = "identity") +
  geom_line(size=1, alpha = 0.9, stat = "identity") +
  scale_color_manual(values = pal_cpalfac) +
  facet_wrap(~crime_title) +
  labs(
    title = "Select Group A Incidents Over Time",
    subtitle = paste("Between January 2018 -", month(last30, label = TRUE, abbr = FALSE), year(last30)),
    x = "",
    y = "",
    color = 'VARIABLE'
  ) +
  theme_cpal(base_size = 28) +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position = "none")

```

## Arrest in the Last Month
```{r, Arrests, echo = FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=10, layout = "l-body-outset"}
arrest_table <- dpdMonth %>%
  st_drop_geometry(.) %>%
  group_by(ArOccur, crime_title) %>%
  summarize(count = n()) %>%
  pivot_wider(names_from = ArOccur, values_from = count) %>%
  mutate(NO = ifelse(is.na(NO), 0, NO),
         YES = ifelse(is.na(YES), 0, YES),
         TOTAL = YES+NO,
         ArPercent = paste0(round(YES/TOTAL, digits = 3)*100, "%")) %>%
  select(crime_title, YES, TOTAL, ArPercent)

reactable(arrest_table,
          compact = TRUE,
          class = "cpal-tbl",
          columns = list(
            crime_title = colDef(
              name = "NIBRS"),
            YES = colDef(
              name = "Total Arrests"),
            TOTAL = colDef(
              name = "Total Incidents"),
            ArPercent = colDef(
              name = "Arrest Percentage")
            ),
          theme = reactableTheme(
      headerStyle = list(
        "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
        "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
        borderColor = "#555"
        )
      )
            )
```

## Incidents by Time Period {-}

Facet wrap pie charts of incident types grouped by time of day (early morning, morning, afternoon, evening, late night) 

```{r, Facet Pie Charts, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
dpdIncidents %>%
  filter(Year == thisyear & Month == thismonth) %>%
  st_drop_geometry(.) %>%
  mutate(ToD = factor(ToD, levels = c("Early Morning", "Morning", "Afternoon", "Evening", "Night"))) %>%
  group_by(crime_title, ToD) %>%
  summarize(count = n()) %>%
  ungroup(.) %>%
  group_by(crime_title) %>%
  mutate(perc = count/sum(count)) %>%
  ggplot(aes(x = ToD, y=perc, fill = crime_title)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::percent(perc, accuracy = 0.1)),
            vjust = 2, color = "#FFFFFF", fontface = "bold", size = 7) +

  facet_wrap(~crime_title) +
  scale_fill_manual(values = pal_cpalfac) +
  labs(
    title = "Incidents by Time Period",
    x = "",
    y = "",
    color = 'VARIABLE'
  ) +
  theme_cpal(base_size = 28) +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position = "none")


```

## Incidents by Location of Incident {-}

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
dpdIncidents %>%
  filter(Year == thisyear & Month == thismonth) %>%
  st_drop_geometry(.) %>%
  group_by(premise) %>%
  summarize(victim_count = n()) %>%
  arrange(desc(victim_count)) %>%
  mutate(perc = victim_count/sum(victim_count)) %>%
  slice_head(., n = 7) %>%
  ggplot(
    aes(
      x = premise,
      y = victim_count,
      #      group = crime_category,
      #      color = comprace,
      fill = premise,
    )
  ) +
  geom_bar(size=1, alpha = 1, stat = "identity") +
  scale_fill_manual(values = pal_cpalfac) +
  geom_text(aes(label = scales::percent(perc, accuracy = 0.1)),
            vjust = 2, color = "#FFFFFF", fontface = "bold", size = 7) +
  #  scale_size_manual(values = 3) +
  #  theme(legend.position = "none") +
  labs(
    title = "Violent Gun Crime by Premise",
    subtitle = paste("For the month of", month(last30, label = TRUE, abbr = FALSE), year(last30)),
    x = "",
    y = "",
    color = 'VARIABLE'
  ) +
  theme_cpal(base_size = 28) +
  theme(axis.text.x=element_text(angle=30, hjust=0.7, vjust = 0.7),
        legend.position = "none")

## Incidents Within Multi-Family Residences {-}

#Table or bar chart describing the most common types of incidents within multi-family premises (include parking lots, buildings, etc)

## Incidents Within Single-Family Residences {-}

#Table or bar chart describing the most common types of incidents within single-family premises (include occupied, unoccupied, etc)

## Incidents Within Commercial Businesses {-}

#Table or bar chart describing the most common types of incidents within commercial premises (include parking lots, tabc location, occupied/vacant, etc)


```

## Incidents by Race of Victim {-}

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.width=10,fig.height=8}
plot.demo <- as_tibble(cbind(comprace = c("American Indian or Alaska Native", "Asian","Black", "Hispanic or Latino", "White"),
                             type = c("City", "City", "City", "City", "City"),
                             per = c(0.014, 0.0336, 0.2395, 0.4180, 0.2897))) %>%
  mutate(per = as.double(per))

plot.vgc <- dpdIncidents %>%
  filter(Year == thisyear & Month == thismonth) %>%
  st_drop_geometry(.) %>%
  drop_na(comprace) %>%
  group_by(comprace) %>%
  summarize(victim_count = n()) %>%
  mutate(type = "Incident Victims",
         per = victim_count/sum(victim_count)) %>% #EDIT FOR OTHER GEOGRAPHIES
  filter(comprace != "Unknown",
         comprace != "NH",
         comprace != "Native Hawaiian/Pacific Islander",
         #comprace != "American Indian or Alaska Native",
         comprace != "Middle Eastern",
         comprace != "")

plot.race <- full_join(plot.demo, plot.vgc) %>%
  mutate(type = as.factor(type)) %>%
  filter(per != 0)

plot.race %>%
  ggplot(
    aes(
      x = comprace,
      y = per,
      fill = type
    )
  ) +
  geom_bar(size=1, alpha = 1, stat = 'identity', position = 'dodge') +
  scale_y_continuous(labels = scales::percent, limits = c(0,0.6)) +
  scale_fill_manual(values = pal_cpaltri) +
  theme_cpal(base_size = 28) +
  theme(legend.position = "bottom",
        legend.title=element_blank()) +
  labs(
    title = "Incident Victims by selected incident types",
    subtitle = "In comparison to city demographics (%).",
    x = "",
    y = "",
    color = ""
  )

```

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.width=12,fig.height=8, layout="l-body-outset", include = FALSE}
## Select Group A Incidents by Division {-}

incident_nibrs_div <- dpdIncidents %>%
  mutate(division = str_to_title(division)) %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month, division, crime) %>%
  summarize(count = n())

incident_vgc_div <- dpdIncidents %>%
  mutate(division = str_to_title(division)) %>%
  filter(vgc_flag == 1) %>%
  mutate(crime = "VIOLENT GUN CRIME") %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month, division, crime) %>%
  summarize(count = n())

incident_table_div <- full_join(incident_nibrs_div, incident_vgc_div) %>%
  filter(Year > 2016) %>%
  mutate(crime_title = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", "Murder & Nonnegligent Manslaughter",
                              ifelse(crime == "VIOLENT GUN CRIME", "Violent Gun Crime",
                                     ifelse(crime == "AGG ASSAULT - NFV", "Aggravated Assault (NFV)",
                                            ifelse(crime == "ROBBERY-BUSINESS", "Robbery (Business)",
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", "Robbery (Individual)",
                                                          ifelse(crime == "BURGLARY-RESIDENCE", "Burglary (Residence)",
                                                                 ifelse(crime == "BURGLARY-BUSINESS", "Burglary (Business)",
                                                                        "ERROR"
                                                                 ))))))))

report_table_div <- incident_table_div %>%
  filter(Month == recent_month) %>%
  ungroup(.) %>%
  select(-Month) %>%
  pivot_wider(names_from = Year,
              names_prefix = "dpd_",
              values_from = count) %>%
  mutate(dpd_2022 = ifelse(is.na(dpd_2022), 0, dpd_2022),
         dpd_2021 = ifelse(is.na(dpd_2021), 0, dpd_2021), 
         dpd_2020 = ifelse(is.na(dpd_2020), 0, dpd_2020),
         dpd_2019 = ifelse(is.na(dpd_2019), 0, dpd_2019),
         poissonYtY = round(2*(sqrt(dpd_2022)-sqrt(dpd_2021)), digits = 2),
         Year2Year = paste0(round(((dpd_2022-dpd_2021)/dpd_2021)*100, digits = 1), "%"),
         Avg3Year = round((dpd_2020+dpd_2021+dpd_2022)/3, digits = 1),
         Diff3Year = paste0(round(((dpd_2022-Avg3Year)/Avg3Year)*100, digits = 1), "%"),
         crime = factor(crime, levels = c("MURDER & NONNEGLIGENT MANSLAUGHTER", "VIOLENT GUN CRIME", "AGG ASSAULT - NFV", "ROBBERY-BUSINESS", "ROBBERY-INDIVIDUAL", "BURGLARY-RESIDENCE", "BURGLARY-BUSINESS")),
         crime_title = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", "Murder & Nonnegligent Manslaughter",
                              ifelse(crime == "VIOLENT GUN CRIME", "Violent Gun Crime",
                                     ifelse(crime == "AGG ASSAULT - NFV", "Aggravated Assault (NFV)",
                                            ifelse(crime == "ROBBERY-BUSINESS", "Robbery (Business)",
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", "Robbery (Individual)",
                                                          ifelse(crime == "BURGLARY-RESIDENCE", "Burglary (Residence)",
                                                                 ifelse(crime == "BURGLARY-BUSINESS", "Burglary (Business)",
                                                                        "ERROR"
                                                                 ))))))),
         crime_order = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", 1,
                              ifelse(crime == "VIOLENT GUN CRIME", 2,
                                     ifelse(crime == "AGG ASSAULT - NFV", 3,
                                            ifelse(crime == "ROBBERY-BUSINESS", 4,
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", 5,
                                                          ifelse(crime == "BURGLARY-RESIDENCE", 6,
                                                                 ifelse(crime == "BURGLARY-BUSINESS", 7,
                                                                        "ERROR"
                                                                 )))))))) %>%
  filter(crime_order != "ERROR") %>%
  select(-(dpd_2017:dpd_2020), -crime) %>%
  mutate(Year2Year = ifelse(Year2Year == "NaN%", NA, Year2Year)) %>%
  filter(!is.na(Year2Year)) %>%
  relocate(division, crime_order, crime_title, dpd_2021, dpd_2022, poissonYtY, Year2Year, Avg3Year, Diff3Year)

#####
reactable(report_table_div,
          compact = TRUE,
          class = "cpal-tbl",
          groupBy = "division",
          defaultSorted = c("crime_order"),
          columns = list(
            division = colDef(
              name = "Division"),
            crime_order = colDef(
              name = "Order",
                                   show = FALSE),
                                 crime_title = colDef(
                                   name = "NIBRS"),
                                 dpd_2021 = colDef(
                                   name = "2021",
                                   na = "-"),
                                 dpd_2022 = colDef(
                                   name = "2022",
                                   na = "-"),
                                 poissonYtY = colDef(
                                   name = "Poisson E",
                                   defaultSortOrder = "desc",
                                   align = "center"
                                   ),
            Year2Year = colDef(
              name = "One Year",
              defaultSortOrder = "desc",
              align = "center"
              ),
            Avg3Year = colDef(
              name = "Three Year",
              defaultSortOrder = "desc"),
            Diff3Year = colDef(
              name = "Three Years",
              defaultSortOrder = "desc",
              align = "center"
              )
                               ),
          theme = reactableTheme(
      headerStyle = list(
        "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
        "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
        borderColor = "#555"
        )
      ),
                               columnGroups = list(
                                 colGroup(name = paste(month(last30, label = TRUE, abbr = FALSE), "Incidents"), columns = c("dpd_2021", "dpd_2022")),
                                 colGroup(name = "Percent Change", columns = c("Year2Year", "Diff3Year")),
                                 colGroup(name = "Average", columns = c("Avg3Year"))
                               )
)
```